\documentclass[12pt]{article}
\usepackage{achemso}
\usepackage[margin=1.0in]{geometry}
\usepackage{setspace}
\usepackage{amsmath}             % for equation typesetting
\usepackage{amssymb}             % for equation typesetting
\usepackage{mathrsfs} 
\usepackage{wasysym}             % for geometric shapes
\usepackage{color}               % for colored fonts
\usepackage{setspace}            % for 1.5 and double spacing
\usepackage{graphicx}            % main graphics package
\usepackage{wrapfig}             % allow text wrapping around figures
\doublespacing

\usepackage{appendix}

\usepackage[capitalize]{cleveref}
\crefname{figure}{Fig.}{Figs.}
\Crefname{figure}{Figure}{Figures}
\crefname{table}{Tab.}{Tabs.}
\Crefname{table}{Table}{Tables}
\crefname{equation}{Eq.}{Eqs.}
\Crefname{equation}{Equation}{Equations}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}
\crefname{appsec}{appendix}{appendices}
\Crefname{appsec}{Appendix}{Appendixes}

% Mathematical Shortcuts
\newcommand{\pfrac}[2]{\frac{\partial #1}{\partial #2}}   % partial derivative
\newcommand{\difrac}[2]{\frac{d #1}{d #2}}                % derivative
\newcommand{\bpar}[1]{\left( #1 \right)}                  % big parentheses
\newcommand{\bbra}[1]{\left[ #1 \right]}                  % big brackets
\newcommand{\bbar}[1]{\left| #1 \right|}                  % big bars
\newcommand{\bra}[1]{\left\langle #1 \right\vert}         % bra
\newcommand{\ket}[1]{\left\vert #1 \right\rangle}         % ket
\newcommand{\inner}[2]{\left\langle #1 \left\vert\right. #2 \right\rangle}            % bracket
\newcommand{\innerop}[3]{\left\langle #1 \left\vert #2 \right\vert #3 \right\rangle}  % operator matrix element
\newcommand{\innersub}[4]{\langle \bd{#1}_{#2}, \bd{#3}_{#4} \rangle}                 % bracket with subscripts
\newcommand{\half}{\frac{1}{2}}                           % 1/2
\newcommand{\powfrac}[3]{\bpar{\frac{#1}{#2}}^{#3}}       % fraction raised to a power
\newcommand{\ii}{\infty}                                  % infinity symbol
\newcommand{\tquad}{\quad\quad\quad}                      % triple-quad spacing
\renewcommand{\Im}{\text{Im}}                             % imaginary symbol
\renewcommand{\Re}{\text{Re}}                             % real symbol

\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\newcommand*\suchthat[0]{\text{ }\vert\text{ }}
\newcommand*\complexmatfield[1]{\mathbb{C}^{#1\text{x}#1}}
\newcommand*\speciallinear[2]{\mathrm{UnL}(#1,#2)}
\newcommand*\complexspeciallinear[1]{\speciallinear{#1}{\mathbb{C}}}
\newcommand*\generallinear[2]{\mathrm{GL}(#1,#2)}
\newcommand*\complexgenerallinear[1]{\generallinear{#1}{\mathbb{C}}}
\newcommand*\mat[1]{\boldsymbol{#1}}

\newcommand*\vc[1]{\boldsymbol{#1}}


\title{General Exam}
\date{October 19, 2016 \\ CHB 239}
\author{David Williams-Young\\ Department of Chemistry, University of Washington}


\begin{document}
\maketitle


\newpage
\section{Introduction}

Electronic and nuclear motion are fundamental to our understanding of chemical
and physical phenomena. From electronic reorganization upon photo-induced charge
transfer in the chromophores of solar cells to the ultrafast excitonic dynamics
in nanocrystalline materials, molecular dynamics lies at the heart of chemistry.
As such, to properly model these phenomena theoretically, we must often venture
into the time-domain to capture the physics necessary to completely understand
the problem at hand. As such, this task constitutes a primary focus of modern
\emph{ab initio} quantum molecular theory. 

In the quantum mechanical description of molecular systems, neglecting explicit
coupling to a quantized photon field, time evolution of the total molecular wave
function, $\ket{\Psi (t)}$ is governed by the Hamiltonian wave equation 
(in atomic units),
\begin{equation}
\mathscr{H}(t) \ket{\Psi (t)} = i\partial_t \ket{\Psi(t)} \quad,
\label{eq:WaveEq}
\end{equation}
where $\mathscr{H}(t)$ is the time--dependent Hamiltonian, $\partial_t$ is a
partial derivative with respect to time, and $t$ is a time parameter. We have
adopted the standard description of the wave function as a state vector rather
than an explicit function of coordinates. The moieties enclosed in parentheses
are taken to be parameters, i.e. the wave function is parameterized by time.  In
principle, one may solve (in some approximate manner) \cref{eq:WaveEq}
simultaneously for both the electronic and nuclear degrees of freedom
explicitly. This approach is, however, intractable for quantum systems exceeding
more than a few particles. To simplify the solutions of \cref{eq:WaveEq}, one
may formally decompose $\ket{\Psi (t)}$ into a product of nuclear and electronic
wave functions,
\begin{equation} 
\ket{\Psi (t)} = \ket{\Phi(\vc{R}(t),t)}\otimes\ket{\Theta(t)} 
\quad .  
\label{eq:exactSepElecNuc}
\end{equation} 
%Here, $\vc{r}$ and $\vc{R}$ are the electronic and nuclear coordinates
%respectively, $t$ is a time parameter, and $\Phi(\vc{r} \vert \vc{R}, t)$ and
%$\Theta(\vc{r} \vert t)$ are the time--dependent electronic and nuclear wave
%functions respectively. The vertical separator found in the parameters of the
%above wave functions are used to denote a parametric time dependence of the
%function on those parameters that appear to the right of the separator. 
Here, $\ket{\Phi(\vc{R}(t),t)}$ and $\ket{\Theta (t)}$ are the electronic and
nuclear wave functions respectively and $\vc{R}(t)$ is the expectation value of
the nuclear position operator, $\vc{R}(t) =
\innerop{\Theta(t)}{\hat{\vc{R}}}{\Theta(t)}$.  It should be noted that the
separation of the total molecular wave function in \cref{eq:exactSepElecNuc} is
formally exact\cite{Gross10_PRL123002, Cederbaum08_JCP124101} without loss of
generality in the non-relativistic treatment of quantum mechanics. Even so, the
equations of motion for the time evolution of \cref{eq:exactSepElecNuc} are
wildly complicated\cite{Ghosh15_MP1} and would only be practical for molecules
consisting of a few atoms.  Nevertheless, this explicitly exact treatment allows
one to, at least in in principle, bifurcate the treatment of quantum molecular
dynamics into explicit treatment of the electronic and nuclear time dependence
separately.  Thus, in order to theoretically treat molecular dynamics
accurately, one must be able to treat both the electronic and nuclear degrees of
freedom to some level of accuracy.  This separation will be the underlying
principle of the future directions that I propose in this work.

Over the years, perhaps the most successful \emph{ab initio} treatment of the
electronic degrees of freedom of the total wave function has been in the
configuration interaction (CI) expansion of the electronic wave function. In the
CI expansion, the exact electronic wave function is written as a linear
combination of Slater determinants,
\begin{equation}
\ket{\Phi (\vc{R}(t),t)} = \sum_{\alpha \in \{\phi_p \}}  
  \ket{\psi_\alpha (\vc{R}(t),t)} c_\alpha(t)
\quad ,
\label{eq:CIExp}
\end{equation}
where $\left\lbrace\ket{\psi_\alpha}\right\rbrace$ is a set of Slater
determinantal electronic wave functions, $\left\lbrace c_\alpha \right\rbrace$
is a set of expansion coefficients, and $\left\lbrace \phi_p \right\rbrace$ is a
set of single--particle Fock states, hereafter referred to as \emph{molecular
orbitals} (MO).  In principle, the above expansion is exact given that all
possible electronic configurations are included. Given that each of the Slater
determinants is written as a linear combination of MOs, there are a total of
$\mathcal{C}(N,K)$ possible determinants that would need to be included in the
expansion, where $N$ and $K$ are the total number of electrons and the
cardinaliy of $\left\lbrace \phi_p \right\rbrace$, respectively. Such an
exhaustive expansion is called \emph{full configuration interaction} (FCI) and
is rendered impractical for the majority of systems due to the rapid growth of
$\mathcal{C}(N,K)$.

In practice, we  must restrict which Slater determinants we include in
\cref{eq:CIExp}. As a result, a vast proportion of research in the field of
electronic structure theory has been centered on the proper way to restrict the
expansion in such a way that all of the relevant physics is maintained while
extraneous information may be discarded. Such methods as complete active space
(CAS), multireference configuration interaction (MRCI), density matrix
renormalization group (DMRG), particle--particle propagator methods , and
truncated CI methods may all be described, at their core, as algorithms to
provide such a filtering of the physically relevant determinants of the CI
expansion. While each of these methods have their faults and accolades, whether
it be accuracy or computational burden, the future directions proposed in this
work will primarily focus on CAS and MRCI methods, which will be explained
further in \cref{sec:Future}, though the particle--particle propagator methods
will be discussed to some extent in \cref{sec:pp-X2C,sec:pp-TSH}.

To properly treat the nuclear degrees of freedom, there exist a vast plethora
of computational methods to treat the nuclear dynamics to varying degrees of
accuracy. In the proposed work, I will restrict the discussion of these methods
to only include those that are semi-classical. Semi-classical nuclear dynamics
algorithms are those that treat the nuclear motion classically (by some variant
of Newton's equations of motion) while maintaining a quantum description of the
electronic degrees of freedom. That is to say that we may write the nuclear 
wave function as a product of delta functions centered at the classical nuclear
positions,
\begin{equation}
\inner{\vc{R}}{\Theta (t)} = 
  \prod_{A = 0}^{N_\mathrm{atoms}} \delta^3(\vc{R} - \vc{R}_A(t))
  \quad,
  \label{eq:ClassicalNuclei}
\end{equation}
where $N_\mathrm{atoms}$ is the number of nuclei in the molecule.  This is
often a valid approach, as the nuclei of molecules are orders of magnitude
heavier than the electrons, and thus move much slower. That is not to say that
nuclei are not, strictly speaking, quantized in their time evolution, but
rather that a large proportion of quantum effects in full molecular dynamics
are due to the time dependence of the electrons. Thus it is usually a valid
assumption that the nuclei time evolve classically relative to their electronic
counterparts. That being said, there are cases in which the quantum evolution
of the nuclei must be explicitly treated, such as tunneling in lighter atoms,
in which these semi-classical methods will not suffice to capture the necessary
physics.

Within the semi-classical treatment of molecular dynamics, computational methods
may be further classified as either adiabatic or non-adiabatic. Both of these
classifications may be characterized by their treatment of the transversal of
the manifold of energy eigenstates,
\begin{equation}
\mathscr{H}(t) \ket{\Psi_I (t)} = E_I(t) \ket{\Psi_I (t)}
\quad \forall t.
\label{eq:EnergyEig}
\end{equation}
Here the index $I$ labels a particular energy eigenstate while $E_I$ represents
the energy eigenvalue of that state. We will assume that $E_I(t)$ is a smooth
function of $t$ for the sake simplicity. Adiabatic time evolution is rooted in
the adiabatic theorem in that, given an initial state that exists as an energy
eigenstate of $\mathscr{H}(0)$, $\ket{\Psi_I(0)}$, at some time $t' > 0$ in the
future, the quantum system will remain in an eigenstate of the
\emph{corresponding} eigenstate of $\mathscr{H}(t')$, $\ket{\Psi_I(t')}$. In
order for this approximation to be valid (although it can be proven that this
assumption is a best an approximation while an exact adiabatic time evolution
cannot exist unless the quantum system does not time-evolve), it is a
necessary and sufficient condition that the individual energy eigenstates do
not couple though the time-derivative operator,
\begin{equation}
\innerop{\Psi_I(t)}{\partial_t}{\Psi_J (t)}
  \stackrel{\mathrm{\mbox{\footnotesize adiabatic}}}{=} 0
  \quad \forall I,J,t.
  \label{eq:AdiabaticNAC}
\end{equation}
\Cref{eq:AdiabaticNAC} may be interpreted as an inability to transport
information between energy eigenstates within the adiabatic approximation and
thus the inability to leave said eigenstate throughout the time--evolution.  In
the field of \emph{ab initio} semi-classical dynamics, the workhorses of
adiabatic time evolution are Born--Oppenheimer  (BOMD) and Car--Parinello (CPMD)
molecular dynamics. In the proposed work, I will primarily focus on the BOMD
method as a choice for adiabatic time--evolution which will be discussed in
\cref{sec:Future}.  As it should logically follow, non-adiabatic dynamics may be
characterized by the full inclusion of terms like the left-hand side of
\cref{eq:AdiabaticNAC}, i.e.  it is in general not possible or practical to
ignore the coupling between energy eigenstates. As such it is possible to
transverse the manifold of energy eigenstates throughout the time evolution.
In this proposed work, I will limit the discussion of non--adiabatic dynamics to
that of the trajectory surface hopping (TSH) method which will be discussed at
some length in \cref{sec:pp-TSH,sec:Future}.

Recent years have also seen new developments in the realm of relativistic
quantum chemistry.  Relativistic effects, while often neglected in most
standard treatment of electronic structure, can have profound consequences in
chemical systems.\cite{Pyykko12_45} Scalar relativistic effects cause the
contraction of the core electron shells of heavy atoms, but perhaps of even
more consequence is the introduction of spin couplings in the Hamiltonian.
Spin-spin and spin-orbit interactions can affect the electronic spin dynamics
even in light atoms, and a direct consequence of these couplings on excited
states is the loss of degeneracies of spin-eigenstates, giving rise to fine
structure splittings (FSS) in atoms and molecules with symmetry-induced
degeneracies. This degeneracy breaking allows for eigenstates of
\cref{eq:EnergyEig} with differing spin symmetry to couple in ways that are
inaccessible though non--relativisitc treatments.
%While much work has gone into the relativistic treatment of the
%mean--field description of ground state electronic structure, far less has been
%applied to the treatement of excited states.
\textbf{In this work, I will outline my proposed research to accurately simulate
\emph{ab initio} relativistic molecular dynamics using correlated
multi--configurational wave functions}.  
A brief review of the necessary theory for the two--component relativisitc
Hamiltonian that I will later apply to CAS and MRCI wave functions will be given
in \cref{sec:X2C}. I believe this be the the natural progression from my previous
extension of perturbatively correlated excited state methods to two--component
relativistic Hamiltonians\cite{DBWY16_Accepted1} (\cref{sec:pp-X2C}), as well as
my work in the field of TSH\cite{DBWY16_Submitted1} (\cref{sec:pp-TSH}). 


%This work is organized into four subsequent sections.
%\Cref{sec:X2C} will breifly review the necessary theory for the two--component
%relativistic Hamiltonian that I will later apply to CAS and MRCI wave functions.
%\Cref{sec:pp-X2C} focuses on my recent work of the extension of the
%particle-particle Tamm-Dancoff approximation to relativistic Hamiltonians.
%\cite{DBWY16_Accepted1}
%\Cref{sec:pp-TSH} describes my contribution to semi-classical non-adiabatic
%molecular dynamics through the application of the particle-particle Tamm-Dancoff
%approximation to the trajectory surface hopping method.
%Finally,
%\Cref{sec:Future} outlines my plan for future development in the field of
%electronic structure and dynamics in the relativistic treatment of adiabatic and
%non-adiabatic phenomena in molecular systems.
In the following  I will use indices $i,j,\dots$ to refer to occupied MOs,
$a,b,\dots$ to refer to virtual MOs, and $p,q,\dots$ to refer to any orbital
regardless of occupancy.  All MOs are expanded in terms of real atomic-orbitals
(AO).

%\section{The Chebyshev Expansion of the Quantum Propagator}
%\label{sec:Chebyshev}

\section{The Exact Two-Component Method}
\label{sec:X2C}
In the relativistic treatment of quantum molecular systems, the Hamiltonian
of interest is that of the Dirac Hamiltonian,
\begin{equation}
\mathscr{H}_D = 
\begin{pmatrix}
  V \otimes \vc{I}_2 && c \text { } p^k \otimes \vc{\sigma}_k \\
  c \text { } p^k \otimes \vc{\sigma}_k && (V - 2mc^2) \otimes \vc{I}_2
\end{pmatrix} \quad ,
\label{eq:DiracHam}
\end{equation}
which describes the relativistic behaviour of a single electron (fermion) in the
presence of a scalar potential $V$. Here, $V$ collects all scalar potential
terms, $\vec{p}$ is the momentum operator (acting as an internal vector
potential), $\vec{\vc{\sigma}}$ is a vector whose elements are the Pauli
matricies, and $\vc{I}_2$ is the 2x2 identity matrix. The implicit parametric
depencence on time has been dropped for brevity. The fact that this is a single
particle operator is quite concequential in the relativistic treatment of
molecules in that $\mathscr{H}_D$ describes the mean-field quantum nature of a
single electron in the presence of the $N-1$ other electrons and the nuclei
(encapsulated in $V$). This is due to the fact that, in general, a Lorentz
invariant many--body Dirac equation does not exist if the Coulomb interaction is
taken to be the interaction potential between charges particles. In this work I
will not be explicitly taking into account relativistic retardation effects,
hence this treatment is not completely relativistic. This treatment does,
however, contain the chemically relevant relativistic effects such as
spin--orbit and spin--spin coupling as well as scalar relativistic effects.

From the matrix representation of $\mathscr{H}_D$, it is clear to see that it
acts upon a four--dimensional Hilbert space known as a bispinor field ,
$\ket{\Phi^{4c}}$, whos components are themselves two--dimensional,
\begin{equation}
\ket{\Phi^{4c}} = \begin{pmatrix}
 \ket{\Phi_L} \\ \ket{\Phi_S}
\end{pmatrix} \quad.
\end{equation}
Here, $\ket{\Phi_L}$ and $\ket{\Phi_S}$ are the so--called large and small
components of the bispinor respectively, and are both two--dimensional in the
spin manifold (i.e. spinors).

Although it is possible, in principle, to obtain the eigenstates of 
\cref{eq:DiracHam} directly, it as often advantageous from a practical as well
as aesthetic perspective to transfrom the full four component relativistic
equations into a decoupled two component form. This is often convientient as the
resulting expressions closely resemble those found in standard electronic
structure theory and therefore allow the employment of standard electronic
structure methods with only minor modifications. In general, such a
transformation takes the form of a unitary operator, $\mathscr{U}$, such that
\begin{equation}
\mathscr{U}: 
\mathscr{H}_D \mapsto \begin{pmatrix}
\mathscr{H}_+ && \vc{0}_2 \\ \vc{0}_2 && \mathscr{H}_- 
\end{pmatrix} \quad \text{ and } \quad
\ket{\Phi^{4c}} \mapsto \begin{pmatrix}
 \ket{\Phi^{2c}} \\ \vc{0}_2
\end{pmatrix} \quad.
\end{equation}
Thus making the transformed $\ket{\Phi^{2c}}$ an eigenstate of $\mathscr{H}_+$.
The transformation under $\mathscr{U}$, in effect, folds the contributions from
the small component of the wavefuction into the large components. In principle,
such a transformation is exact if a proper $U$ may be found. It is the case
that such an exact transformation is not practical to obtain due to the
effective many--body $V$, thus approximate decoupling schemes must be developed.
Several decoupling schemes have been explored in recent years, but in this work
I will only emply the use of the ``exact" two--component (X2C) method, for which
the exact details may be found elsewhere.

Using the X2C transformation method, we are able to obtain a total effecting
two--component Hamiltonian, which may be written in second quantized form,
\begin{equation}
\mathscr{H}_\mathrm{X2C} = h_{pq}^\mathrm{X2C}c_p^\dagger c_q + 
  \frac{1}{2} \inner{pq}{rs} c_p^\dagger c_r^\dagger c_s c_q + V_\mathrm{NN}
  \label{eq:X2CHam}
\end{equation}
where $c_p$ and $c_p^\dagger$ are single spinor particle annihilation and
creation operators respectively. $\vc{h}^\mathrm{X2C}$ is the effective
relativistic core--Hamiltonian which includes both spin--orbit and scalar
relativistic effects, $\inner{\cdot}{\cdot}$, is the Coulomb operator expressed
in the basis of spinor MOs and $V_\mathrm{NN}$ is the nuclear--nuclear repulsion
energy. Expressing \cref{eq:X2CHam} in basis of a single Slater determinant and
expanding each spinor MO in the AO basis ($\lbrace \chi_\mu \rbrace$),
\begin{equation}
\phi_p ( \vc{r} ) = \sum_\mu \chi_\mu(\vc{r}) C_{\mu p}
\qquad
C_{\mu p} = \begin{pmatrix}
C_{\mu p}^\alpha \\ C_{\mu p}^\beta
\end{pmatrix}
\quad , \label{eq:AO2MO}
\end{equation}
we arrive at a relativisitc two--component analouge of the Roothaan--Hall
equations familiar to traditional mean--field electronic structore theory,
\begin{align}
&\frac{1}{2}\left(
  F^S_{\mu\nu} \cdot \vc{I}_2 + F^k_{\mu\nu} \cdot \vc{\sigma_k}
\right) C_{\nu p} = (S_{\mu\nu}\cdot \vc{I}_2) C_{\nu p}
\epsilon_p\label{eq:Roothaan}\\
\nonumber \\
&F^S_{\mu\nu} = \left(
  2\left( \mu\nu \vert \kappa\lambda \right) - 
  \left(  \mu\lambda \vert \kappa\nu \right) 
\right) P^S_{\lambda \kappa} \nonumber \\
&F^k_{\mu\nu} = -\left(  \mu\lambda \vert \kappa\nu \right) P^k_{\lambda \kappa}\\
\nonumber \\
&\vc{P}^S = \frac{1}{2}\mathrm{Tr}_\sigma (\vc{P}\vc{I}_2) \nonumber \\
&\vc{P}^k = \frac{1}{2}\mathrm{Tr}_\sigma (\vc{P}\vc{\sigma}_k)
\label{eq:traceRel}\\
&P_{\mu\nu} = C_{\mu i} C_{\nu i}^\dagger \nonumber
\end{align}
Here, $\vc{S}$ is the AO overlap matrix, and $(\cdot \vert \cdot)$ is the
Coulomb operator in the AO basis. $\{\epsilon_p\}$ is the set of MO
eigenenergies. $\vc{F}^S$,$\vc{F}^k$ and $\vc{P}^S,\vc{P}^k$ are the scalar and
vector parts of the spinor Fock and density matricies in the AO basis
respectively, and are related to the full spinor operators in terms of the spin
trace, $\mathrm{Tr}_\sigma$, relations of \cref{eq:traceRel}. Is important to
note that $C_{\mu p} \in \mathbb{C}^2$ (where the $\alpha$ and $\beta$ incidies
represent inseperable spin-up and spin-down components of $C_{\mu p}$), thus the
total spinor density (and thus the spinor Fock) are 2x2 complex matricies.
\Cref{eq:Roothaan} may be solved self--consistantly to obtain the lowest energy
(single) Slater determinantal description of the many--body system.
% Relativistic pp-TDA
\section{The Relativistic Particle-Particle Tamm-Dancoff Approximation}
\label{sec:pp-X2C}

The ability to accurately predict and characterize the electronically excited
states of molecular systems is paramount to a complete understanding of many
chemical phenomena. As such, excited states are the central focus of many
fields of physical chemistry, most prominent being that of spectroscopy. Due to
this centralized importance and the need to efficiently and accurately predict
excited states properties, much  effort has been devoted over the years towards
the modeling of excitation energies and oscillator strengths.

Recently, the particle-particle random phase approximation (pp-RPA) and
Tamm-Dancoff approximation (pp-TDA), which have been standard trade tools of
the nuclear physics community in the treatment of the many--body correlation
energy for low matter density systems\cite{SchuckBook_04}, have been extended
to the treatment the correlation energy and excitation energies of quantum
molecular systems within a finite basis set.\cite{Yang13_224105,Yang13_18A522,
Yang13_174110,Yang13_104112,Yang13_030501,Yang09_066403,Bulik13_104113}.
Although this introduction into the quantum chemistry community in relatively
recent, a wealth of effort has been afforded to the rigorous investigation of
these methods in a variety of different contexts, involving the evaluation of
excitation energies\cite{Yang13_224105,Yang13_18A522,Yang13_174110}, excited
state properties and geometry optimizations\cite{Yang15_1025}, and the
treatment of non-adiabatic phenomena such as non-adiabatic
dynamics\cite{DBWY16_Submitted1,Liu14_244105}  and the description of conical
intersections\cite{Yang16_2407}.

I have recently extended the pp-TDA to relativistic and non--relativistic
two--component Hamiltonians\cite{DBWY16_Accepted1} to accurately compute the
FSS of a variety of atomic and molecular systems. In this work, I will briefly
review the two--component tranformation for that I will later apply to CAS and
MRCI wavefunctions and further present the FSS results obtained in my previous
work.  

% TSH with pp-TDA
\section{Trajectory Surface Hopping within the Particle-Particle Tamm-Dancoff Approximation}
\label{sec:pp-TSH}
Test\cite{DBWY16_JCTC935,DBWY16_Accepted1,DBWY16_Submitted1,DBWY16_Submitted2}

% Future
\section{Future Directions: Relativistic Electronic Dynamics with Multi-Configurational Wave Functions}
\label{sec:Future}

\bibliography{Journal_Short_Name,ppSH,Li_Group_References,GE,DBWY,Egidi_References}
\end{document}
