\documentclass[12pt]{article}
\usepackage{achemso}
\usepackage[margin=1.0in]{geometry}
\usepackage{setspace}
\usepackage{amsmath}             % for equation typesetting
\usepackage{amssymb}             % for equation typesetting
\usepackage{mathrsfs} 
\usepackage{wasysym}             % for geometric shapes
\usepackage{color}               % for colored fonts
\usepackage{setspace}            % for 1.5 and double spacing
\usepackage{graphicx}            % main graphics package
\usepackage{wrapfig}             % allow text wrapping around figures
\doublespacing

\usepackage{appendix}

\usepackage[capitalize]{cleveref}
\crefname{figure}{Fig.}{Figs.}
\Crefname{figure}{Figure}{Figures}
\crefname{table}{Tab.}{Tabs.}
\Crefname{table}{Table}{Tables}
\crefname{equation}{Eq.}{Eqs.}
\Crefname{equation}{Equation}{Equations}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}
\crefname{appsec}{appendix}{appendices}
\Crefname{appsec}{Appendix}{Appendixes}

% Mathematical Shortcuts
\newcommand{\pfrac}[2]{\frac{\partial #1}{\partial #2}}   % partial derivative
\newcommand{\difrac}[2]{\frac{d #1}{d #2}}                % derivative
\newcommand{\bpar}[1]{\left( #1 \right)}                  % big parentheses
\newcommand{\bbra}[1]{\left[ #1 \right]}                  % big brackets
\newcommand{\bbar}[1]{\left| #1 \right|}                  % big bars
\newcommand{\bra}[1]{\left\langle #1 \right\vert}         % bra
\newcommand{\ket}[1]{\left\vert #1 \right\rangle}         % ket
\newcommand{\inner}[2]{\left\langle #1 \left\vert\right. #2 \right\rangle}            % bracket
\newcommand{\innerop}[3]{\left\langle #1 \left\vert #2 \right\vert #3 \right\rangle}  % operator matrix element
\newcommand{\innersub}[4]{\langle \bd{#1}_{#2}, \bd{#3}_{#4} \rangle}                 % bracket with subscripts
\newcommand{\half}{\frac{1}{2}}                           % 1/2
\newcommand{\powfrac}[3]{\bpar{\frac{#1}{#2}}^{#3}}       % fraction raised to a power
\newcommand{\ii}{\infty}                                  % infinity symbol
\newcommand{\tquad}{\quad\quad\quad}                      % triple-quad spacing
\renewcommand{\Im}{\text{Im}}                             % imaginary symbol
\renewcommand{\Re}{\text{Re}}                             % real symbol

\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\newcommand*\suchthat[0]{\text{ }\vert\text{ }}
\newcommand*\complexmatfield[1]{\mathbb{C}^{#1\text{x}#1}}
\newcommand*\speciallinear[2]{\mathrm{UnL}(#1,#2)}
\newcommand*\complexspeciallinear[1]{\speciallinear{#1}{\mathbb{C}}}
\newcommand*\generallinear[2]{\mathrm{GL}(#1,#2)}
\newcommand*\complexgenerallinear[1]{\generallinear{#1}{\mathbb{C}}}
\newcommand*\mat[1]{\boldsymbol{#1}}

\newcommand*\vc[1]{\boldsymbol{#1}}
\newcommand*\op[1]{\mathcal{#1}}


\title{General Exam}
\date{October 19, 2016 \\ CHB 239}
\author{David Williams-Young\\ Department of Chemistry, University of Washington}


\begin{document}
\maketitle


\newpage
\section{Introduction}

Electronic and nuclear motion are fundamental to our understanding of chemical
and physical phenomena. From  geometric reorganization upon photo-induced charge
transfer in the chromophores of photovoltaics  to vibrational mediation of
intersystem crossing that gives rise to phosphorescence,  molecular dynamics
lies at the heart of chemistry.  As such, to properly model these phenomena
theoretically, we must often venture into the time-domain to capture the physics
necessary to completely understand the problem at hand. As such, this task
constitutes a primary focus of modern \emph{ab initio} quantum molecular theory.
{\bf
Although much work has gone into the field of \emph{ab initio} molecular
dyanmics, the scope of inquiry has been primarily limited to a non--relativistic
regime.
}

Relativistic effects, while often neglected in most standard treatment of
quantum mechanics, can have profound consequences in chemical
systems.\cite{Pyykko12_45} Scalar relativistic effects cause the contraction of
the core electron shells of heavy atoms, but perhaps of even more consequence is
the introduction of spin couplings in the Hamiltonian.  Spin-spin (SSC) and
spin-orbit (SOC) coupling can affect the electronic spin dynamics even in light
atoms. A direct consequence of these couplings on the electronic manifold is the
breaking of spin--symmetry as the Hamiltonian no longer commutes with the spin
operator, $\op{S}$. This breaking of spin--symmetry is what allows, at an
operator level, for formally spin--forbidden processes to occur, namely
intersystem crossing (ISC) and internal conversion. Although some approaches
have been purposed to account for SOC in the electronic manifold
perturbatively\cite{Thiel14_JCP124101}, these approximation will break down
whenever SOC is non--negligible. Hence a variational approach much be adopted to
accrount for these interactions in the general case.  Thus to accurately treat
these inheriently relativistic and dynamical phenomena theoretically, one must
provide an \emph{ab initio} description of the relativistic effects thoughout
the time--evolution of the quantum system.  
{\bf 
In this work, I will outline my proposed research to accurately simulate
\emph{ab initio} relativistic molecular dynamics using correlated wave
functions.
}  

In the quantum mechanical description of molecular systems, neglecting explicit
coupling to a quantized photon field, time evolution of the total molecular wave
function, $\ket{\Psi (t)}$ is governed by the Hamiltonian wave equation 
(in atomic units),
\begin{equation}
\op{H}(t) \ket{\Psi (t)} = i\partial_t \ket{\Psi(t)} \quad,
\label{eq:WaveEq}
\end{equation}
where $\op{H}(t)$ is the time--dependent Hamiltonian, $\partial_t$ is a
partial derivative with respect to time, and $t$ is a time parameter. We have
adopted the standard description of the wave function as a state vector rather
than an explicit function of coordinates. The moieties enclosed in parentheses
are taken to be parameters, i.e. the wave function is parameterized by time.  In
principle, one may solve (in some approximate manner) \cref{eq:WaveEq}
simultaneously for both the electronic and nuclear degrees of freedom
explicitly. This approach is, however, intractable for quantum systems exceeding
more than a few particles. To simplify the solutions of \cref{eq:WaveEq}, one
may formally decompose $\ket{\Psi (t)}$ into a product of nuclear and electronic
wave functions,
\begin{equation} 
\ket{\Psi (t)} = \ket{\Phi(\vc{R}(t),t)}\otimes\ket{\Theta(t)} 
\quad .  
\label{eq:exactSepElecNuc}
\end{equation} 
Here, $\ket{\Phi(\vc{R}(t),t)}$ and $\ket{\Theta (t)}$ are the electronic and
nuclear wave functions respectively and $\vc{R}(t)$ is the expectation value of
the nuclear position operator, $\vc{R}(t) =
\innerop{\Theta(t)}{\hat{\vc{R}}}{\Theta(t)}$.  It should be noted that the
separation of the total molecular wave function in \cref{eq:exactSepElecNuc} is
formally exact\cite{Gross10_PRL123002, Cederbaum08_JCP124101} without loss of
generality in the non-relativistic treatment of quantum mechanics. Even so, the
equations of motion for the time evolution of \cref{eq:exactSepElecNuc} are
wildly complicated\cite{Ghosh15_MP1} and would only be practical for molecules
consisting of a few atoms.  Nevertheless, this explicitly exact treatment allows
one to, at least in in principle, bifurcate the treatment of quantum molecular
dynamics into explicit treatment of the electronic and nuclear time dependence
separately.  Thus, in order to theoretically treat molecular dynamics
accurately, one must be able to treat both the electronic and nuclear degrees of
freedom to some level of accuracy.  This separation will be the underlying
principle of the future directions that I propose in this work.

% Paragraph on the fact that there exsist a shit ton of methods to solve these
% problems but we're going to focus on CAS/MRCI coupled with TSH


I believe that this proposed work is the natural progression from my previous
works in the field the relativistic electronic structure
theory\cite{DBWY16_Accepted1} and nonrelativistic non--adiabatic semiclassical
molecular dynamics.\cite{DBWY16_JCTC935,DBWY16_Submitted1} In the following, I
will present a brief overeview of the necessary theory for treatment of
non--adiabatic semiclassical dynamics (\cref{sec:TSH}) and the relativistic
Hamiltonian (\cref{sec:X2C}) which will be used throughout this work.
In \cref{sec:pp-X2C,sec:pp-TSH}, I will briefly recall my advances in relevant
fields and how they interconnect to my proposed research.  Finally,
\cref{sec:Future} will outline my plan for the advancement of \emph{ab initio}
non-adiabatic dynamics to correlated relativistic wave functions.

%In the following  I will use indices $i,j,\dots$ to refer to occupied MOs,
%$a,b,\dots$ to refer to virtual MOs, and $p,q,\dots$ to refer to any orbital
%regardless of occupancy.  All MOs are expanded in terms of real atomic-orbitals
%(AO). The Einstein summation convention will be assumed throughout.

\section{Trajectory Surface Hopping}
\label{sec:TSH}

When obtaining the time evolution of the total molecular wave function, it is
often advantageous to work in so called adiabatic basis (the eigenstates of
\cref{eq:WaveEq}) of quantum states,
\begin{equation}
\op{H}(t) \ket{\Psi_I (t)} = E_I(t) \ket{\Psi_I (t)}
\quad \forall t, I \in \mathbb{N},
\label{eq:EigSpec}
\end{equation}
where $\ket{\Psi_I}$ and $E_I$ represent the $I$-th adiabatic wave function and
eigenenergie, respoectively. For the purposes of the current discussion, we will
be assuming a discrete eigenspectrum of the Hamiltonian that is smooth over $t$.
Taking the separating of the total wave function in \cref{eq:exactSepElecNuc},
we may rewrite the total wave function as a linear combitnation of adiabatic
states,
\begin{equation}
\ket{\Psi (t)} = c_I \ket{\Phi_I (\vc{R}(t),t)} \otimes \ket{\Theta_I (t)}
\quad ,
\label{eq:AdiaExp}
\end{equation}
where $\{ c_I \}$ is a set of expansion coefficients.  Given the entire
eigenspectrum of \cref{eq:EigSpec}, the expansion in \cref{eq:AdiaExp} is
formally exact, thus the problem now becomes how to properly obtain the adiabtic
electronic and nuclear wave functions.

There exist a vast plethora of computational methods to treat the nuclear
dynamics to varying degrees of accuracy. In the proposed work, I will restrict
the discussion of these methods to only include those that are semi-classical.
Semi-classical nuclear dynamics algorithms are those that treat the nuclear
motion classically (by some variant of Newton's equations of motion) while
maintaining a quantum description of the electronic degrees of freedom. That is
to say that we may write the nuclear wave function as a product of delta
functions centered at the classical nuclear positions,
\begin{equation}
\inner{\vc{R}}{\Theta (t)} = 
  \prod_{A = 0}^{N_\mathrm{atoms}} \delta^3(\vc{R} - \vc{R}_A(t))
  \quad,
  \label{eq:ClassicalNuclei}
\end{equation}
where $N_\mathrm{atoms}$ is the number of nuclei in the molecule.  This is
often a valid approach, as the nuclei of molecules are orders of magnitude
heavier than the electrons, and thus move much slower. That is not to say that
nuclei are not, strictly speaking, quantized in their time evolution, but
rather that a large proportion of quantum effects in full molecular dynamics
are due to the time dependence of the electrons. Thus it is usually a valid
assumption that the nuclei time evolve classically relative to their electronic
counterparts. That being said, there are cases in which the quantum evolution
of the nuclei must be explicitly treated, such as tunneling in lighter atoms,
in which these semi-classical methods will not suffice to capture the necessary
physics.

Within the semi-classical treatment of molecular dynamics, computational methods
may be further classified as either adiabatic or non-adiabatic. Both of these
classifications may be characterized by their treatment of the transversal of
the manifold of the adiabatic states. Adiabatic time evolution is rooted in
the adiabatic theorem in that, given an initial state that exists as an energy
eigenstate of $\mathscr{H}(0)$, $\ket{\Psi_I(0)}$, at some time $t' > 0$ in the
future, the quantum system will remain in an eigenstate of the
\emph{corresponding} eigenstate of $\mathscr{H}(t')$, $\ket{\Psi_I(t')}$. In
order for this approximation to be valid (although it can be proven that this
assumption is a best an approximation while an exact adiabatic time evolution
cannot exist unless the quantum system does not time-evolve), it is a
necessary and sufficient condition that the individual energy eigenstates do
not couple though the time-derivative operator,
\begin{equation}
\innerop{\Psi_I(t)}{\partial_t}{\Psi_J (t)}
  \stackrel{\mathrm{\mbox{\footnotesize adiabatic}}}{=} 0
  \quad \forall I \not = J,t.
  \label{eq:AdiabaticNAC}
\end{equation}
\Cref{eq:AdiabaticNAC} may be interpreted as an inability to transport
information between energy eigenstates within the adiabatic approximation and
thus the inability to leave said eigenstate throughout the time--evolution.  It
follows that non--adiabatic time evolution encompasses those methods which are
able to transport information though the adiabatic manifold through terms such
as \cref{eq:AdiabaticNAC}. By construction, adiabatic time evolution cannot
capture phenomena such as ISC or internal conversion as they are inheriently
dependant on the ability to trasport information through the adiabatic manifold.
Thus, in this work, I will only consider methods which are both semiclassical
and non--adiabatic.

\section{The Exact Two-Component Method}
\label{sec:X2C}

In the relativistic treatment of quantum molecular systems, the Hamiltonian
of interest is that of the Dirac Hamiltonian,
\begin{equation}
\op{H}_D = 
\begin{pmatrix}
  V \otimes \vc{I}_2 && c \text { } p^k \otimes \vc{\sigma}_k \\
  c \text { } p^k \otimes \vc{\sigma}_k && (V - 2mc^2) \otimes \vc{I}_2
\end{pmatrix} \quad ,
\label{eq:DiracHam}
\end{equation}
which describes the relativistic behaviour of a single electron (fermion) in the
presence of a scalar potential $V$. Here, $V$ collects all scalar potential
terms, $\vec{p}$ is the momentum operator (acting as an internal vector
potential), $\vec{\vc{\sigma}}$ is a vector whose elements are the Pauli
matricies, and $\vc{I}_2$ is the 2x2 identity matrix. The implicit parametric
depencence on time has been dropped for brevity. The fact that this is a single
particle operator is quite concequential in the relativistic treatment of
molecules in that $\op{H}_D$ describes the mean-field quantum nature of a
single electron in the presence of the $N-1$ other electrons and the nuclei
(encapsulated in $V$). This is due to the fact that, in general, a Lorentz
invariant many--body Dirac equation does not exist if the Coulomb interaction is
taken to be the interaction potential between charges particles. In this work I
will not be explicitly taking into account relativistic retardation effects,
hence this treatment is not completely relativistic. This treatment does,
however, contain the chemically relevant relativistic effects such as
SOC and  scalar relativistic effects.
From the matrix representation of $\op{H}_D$, it is clear to see that it
acts upon a four--dimensional Hilbert space known as a bispinor field ,
$\ket{\Phi^{4c}}$, whos components are themselves two--dimensional,
\begin{equation}
\ket{\Phi^{4c}} = \begin{pmatrix}
 \ket{\Phi_L} \\ \ket{\Phi_S}
\end{pmatrix} \quad.
\end{equation}
Here, $\ket{\Phi_L}$ and $\ket{\Phi_S}$ are the so--called large and small
components of the bispinor respectively, and are both two--dimensional in the
spin manifold (i.e. spinors).

Although it is possible, in principle, to obtain the eigenstates of
\cref{eq:DiracHam} directly, it as often advantageous from a practical as well
as aesthetic perspective to transfrom the full four component relativistic
equations into a decoupled two component form. This is often convientient as the
resulting expressions closely resemble those found in non--relativistic
electronic structure theory and therefore allow the employment of standard
electronic structure methods with only minor modifications. In general, such a
transformation takes the form of a unitary operator, $\op{U}$, such that
\begin{equation}
\op{U}: 
\op{H}_D \mapsto \begin{pmatrix}
\op{H}_+ && \vc{0}_2 \\ \vc{0}_2 && \op{H}_- 
\end{pmatrix} \quad \text{ and } \quad
\ket{\Phi^{4c}} \mapsto \begin{pmatrix}
 \ket{\Phi^{2c}} \\ \vc{0}_2
\end{pmatrix} \quad.
\end{equation}
Thus making the transformed $\ket{\Phi^{2c}}$ an eigenstate of $\op{H}_+$.
The transformation under $\op{U}$, in effect, folds the contributions from
the small component of the wavefuction into the large components. In principle,
such a transformation is exact if a proper $U$ may be found. It is the case
that such an exact transformation is not practical to obtain due to the
effective many--body $V$, thus approximate decoupling schemes must be developed.
Several decoupling schemes have been explored in recent years, but in this work
I will only emply the use of the ``exact" two--component (X2C) method, for which
the exact details may be found elsewhere.

Using the X2C transformation method, we are able to obtain a total effecting
two--component Hamiltonian, which may be written in second quantized form,
\begin{equation}
\op{H}_\mathrm{X2C} = h_{pq}^\mathrm{X2C}c_p^\dagger c_q + 
  \frac{1}{2} \inner{pq}{rs} c_p^\dagger c_r^\dagger c_s c_q + V_\mathrm{NN}
  \label{eq:X2CHam}
\end{equation}
where $c_p$ and $c_p^\dagger$ are single spinor particle annihilation and
creation operators respectively. $\vc{h}^\mathrm{X2C}$ is the effective
relativistic core--Hamiltonian which includes both spin--orbit and scalar
relativistic effects, $\inner{\cdot}{\cdot}$, is the Coulomb operator expressed
in the basis of spinor MOs and $V_\mathrm{NN}$ is the nuclear--nuclear repulsion
energy. Expressing \cref{eq:X2CHam} in basis of a single Slater determinant and
expanding each spinor MO in the AO basis ($\lbrace \chi_\mu \rbrace$),
\begin{equation}
\phi_p ( \vc{r} ) = \sum_\mu \chi_\mu(\vc{r}) C_{\mu p}
\qquad
C_{\mu p} = \begin{pmatrix}
C_{\mu p}^\alpha \\ C_{\mu p}^\beta
\end{pmatrix}
\quad , \label{eq:AO2MO}
\end{equation}
we arrive at a relativisitc two--component analouge of the Roothaan--Hall
equations familiar to traditional mean--field electronic structore theory,
\begin{align}
&\frac{1}{2}\left(
  F^S_{\mu\nu} \cdot \vc{I}_2 + F^k_{\mu\nu} \cdot \vc{\sigma_k}
\right) C_{\nu p} = (S_{\mu\nu}\cdot \vc{I}_2) C_{\nu p}
\epsilon_p\label{eq:Roothaan}\\
\nonumber \\
&F^S_{\mu\nu} = \left(
  2\left( \mu\nu \vert \kappa\lambda \right) - 
  \left(  \mu\lambda \vert \kappa\nu \right) 
\right) P^S_{\lambda \kappa} \nonumber \\
&F^k_{\mu\nu} = -\left(  \mu\lambda \vert \kappa\nu \right) P^k_{\lambda \kappa}\\
\nonumber \\
&\vc{P}^S = \frac{1}{2}\mathrm{Tr}_\sigma (\vc{P}\vc{I}_2) \nonumber \\
&\vc{P}^k = \frac{1}{2}\mathrm{Tr}_\sigma (\vc{P}\vc{\sigma}_k)
\label{eq:traceRel}\\
&P_{\mu\nu} = C_{\mu i} C_{\nu i}^\dagger \nonumber
\end{align}
Here, $\vc{S}$ is the AO overlap matrix, and $(\cdot \vert \cdot)$ is the
Coulomb operator in the AO basis. $\{\epsilon_p\}$ is the set of MO
eigenenergies. $\vc{F}^S$,$\vc{F}^k$ and $\vc{P}^S,\vc{P}^k$ are the scalar and
vector parts of the spinor Fock and density matricies in the AO basis
respectively, and are related to the full spinor operators in terms of the spin
trace, $\mathrm{Tr}_\sigma$, relations of \cref{eq:traceRel}. Is important to
note that $C_{\mu p} \in \mathbb{C}^2$ (where the $\alpha$ and $\beta$ incidies
represent inseperable spin-up and spin-down components of $C_{\mu p}$), thus the
total spinor density (and thus the spinor Fock) are 2x2 complex matricies.
\Cref{eq:Roothaan} may be solved self--consistantly to obtain the lowest energy
(single) Slater determinantal description of the many--body system.


\section{The Relativistic Particle-Particle Tamm-Dancoff Approximation}
\label{sec:pp-X2C}
\section{Trajectory Surface Hopping within the Particle-Particle Tamm-Dancoff Approximation}
\label{sec:pp-TSH}
\section{Future Directions: Relativistic Electronic Dynamics with Multi-Configurational Wave Functions}
\label{sec:Future}

\bibliography{Journal_Short_Name,ppSH,Li_Group_References,GE,DBWY,Egidi_References}
\end{document}
